Estimating initial state and quark-gluon plasma medium properties\\ using a hybrid model with nucleon substructure\\ calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV

### 2018-07-16 paper/substructure.tex
Wrapping up experimental observables section

@@

\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
\newcommand{\nch}{N_\text{ch}}
\newcommand{\sqrts}{\sqrt{s_{NN}}}
\newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
\newcommand{\X}{\chi_\mathrm{struct}}
\newcommand{\taufs}{\tau_\mathrm{fs}}
\newcommand{\dmin}{d_\mathrm{min}}
\newcommand{\T}{\tilde{T}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\vnk}[2]{v_#1\{#2\}}
\newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
\newcommand{\order}[1]{$\mathcal O(10^{#1})$}
\newcommand{\note}{\textcolor{theblue}{[?]}}

@@

Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass and set of Woods-Saxon coefficients \cite{MOLLER1995185, DEVRIES1987495}, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to simulation predictions, outweighed by large uncertainties in modeling initial energy deposition and off-equilibrium dynamics \cite{Niemi:2014lha, Song:2011hk, Retinskaya:2013gca, Liu:2015nwa, Kurkela:2016vts}.
The geometry of light ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ signficantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
These nucleon substructure properties are difficult to measure and calculate from first principles and hence contribute significant uncertainty to model predictions of small systems.

@@

Within a picture of independent pairwise collisions between the constituents, a Glauber model model may be used to calculate the probability $P_\mathrm{coll}$ that the two protons collide inelastically at impact parameter $b$. In the limit when the number of constituents $n_c \gg 1$, it yields the particularly simple form
\begin{equation}
  \label{eq:pcoll}
  P_\mathrm{coll} = 1 - \exp[-\sigma_\mathrm{eff}\, T_{pp}(b)], \\
\end{equation}
where $\sigma_\mathrm{eff}$ is an effective cross section for pairwise inelastic collisions between the constituents, and $\sigma_{pp}^\mathrm{inel}$ is the total inelastic proton-proton cross section:
\begin{equation}
  \label{eq:sigma_nn}
  \sigma_{pp}^\mathrm{inel} = \int 2 \pi b\, db\, P_\mathrm{coll}(b).
\end{equation}
The proton densities $\rho_{A,B}$ in Eqn.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eqn.~\eqref{eq:pcoll}.
Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the proton size is comparable to the size of the produced QGP medium.

@@

In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom are uncertain as depicted in Fig.~\ref{fig:substructure}.
We model each nucleon's density $\rho_{A,B}$ as a sum of $n_c$ independent constituents
\begin{equation}
  \label{eq:rho}
  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\mathbf{x} - \mathbf{x_i}),
\end{equation}
where each constituent density $\rho_c$ is described by a Gaussian distribution of width $v$
\begin{equation}
  \label{eq:constituent_density}
  \rho_c(\mathbf{x}) = \frac{1}{(2\pi v^2)^{3/2}} \exp\left(-\frac{\x^2}{2 v^2}\right),
\end{equation}
and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution of width $r$.

@@

A simple remedy is to replace the arithmetic mean in the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
\end{equation}
based on a family of functions known as the generalized mean(s).
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto
  \begin{cases}
    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(T_A, T_B) & p \rightarrow -\infty.
  \end{cases}
\end{equation}
Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} saturation theory calculations, such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.

@@

We thus follow the procedure of the authors and reinterpret our initial transverse entropy density at midrapidity $s(\mathbf{x_\perp}, \tau_0)\vert_{\eta=0}$ as a transverse density of noninteracting massless partons $n(\mathbf{x}, \tau_0)\vert_{\eta=0}$.
The free-streamed energy-momentum tensor $T^{\mu\nu}(x, y, \eta, \tau)$ at transverse coordinate $(x, y)$, midrapidity $\eta=0$, and proper time $\tau > \tau_0$ is then given by
\begin{multline}
  \label{energy-momentum}
  T^{\mu\nu}(x, y, 0, \tau) = \frac{1}{\tau} \int_{0}^{2\pi} d\phi\, n(x - \Delta \tau \cos \phi, y - \Delta \tau \sin \phi) \\ \times
  \begin{bmatrix}
    1 & \cos\phi & \sin\phi  & 0\\
    \cos\phi & \cos^2\phi & \cos\phi\sin\phi & 0 \\
    \sin\phi & \sin\phi \cos\phi & \sin^2\phi & 0\\
    0 & 0 & 0 & 0
  \end{bmatrix},
\end{multline}
where $\Delta\tau$ is the free-streaming time $\Delta\tau = \tau - \tau_0$.
The solution \eqref{energy-momentum} is decomposed in hydrodynamic form
\begin{equation}
  \label{hydro-eqn}
  T^{\mu\nu} = e u^\mu u^\nu - (P + \Pi) \Delta^{\mu\nu} + \pi^{\mu\nu}
\end{equation}
where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} \equiv g^{\mu\nu} - u^\mu u^\nu}$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue,
\begin{equation}
  T^{\mu\nu} u_\nu = e u^\mu.
\end{equation}
The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eqn.~\eqref{hydro-eqn}
\begin{align}
  \Pi &= -\frac{1}{3} \mathrm{Tr}(\Delta_{\mu\nu} T^{\mu\nu}) - P,\\
  \pi^{\mu\nu} &= T^{\mu\nu} - e u^\mu u^\nu + (P + \Pi) \Delta^{\mu\nu}.
\end{align}

@@

The hydrodynamic equations of motion are necessarily closed using an equation of state (EoS) to relate the energy density $e$ and pressure $P$ of the produced medium.
We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval
$165 \le T \le 200$~MeV.
For the lattice EoS, we use a calculation by the HotQCD collaboration for (2+1)-flavor QCD which was extrapolated to the continuum limit \cite{Bazavov:2014pvz}. Recent developments in lattice QCD now enable calculations in (2+1+1)-flavors \cite{Borsanyi:2016ksw}, i.e.\ with thermalized charm quarks, and the additional charm flavor has been shown to visibly affect predictions of $p_T$-differential flow observables \cite{Noronha-Hostler:2018zxc}.
Investigating this sensitivity would thus be a natural target for future improvements to the present work.

@@

For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Tbl.~\ref{tab:design}.
We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
Several of the model parameters require special care and are reparametrized accordingly:
\begin{itemize}[leftmargin=2\parindent]
  \item The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
    The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
    We therefore reparametrize the constituent fluctuations using the standard deviation of the effective nucleon fluctuations
    \begin{equation}
      \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
    \end{equation}
    where $k$ is the shape parameter of the gamma distribution used to fluctuate each constituent.
  \item In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
    \begin{enumerate*}[label=(\roman*)]
      \item a parameter $v$ for the Gaussian width of each constituent,
      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
    \end{enumerate*}
    In the limit $n_c \to \infty$, the composite nucleon's root mean square (rms) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
  \begin{equation}
    \lim_{n_c \to\infty} \mathrm{RMS}\{\rho(\mathbf{x})\} = \sqrt{r^2 + v^2}.
  \end{equation}
  We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
  \begin{equation}
    \label{eq:nucleon_width}
    w = \sqrt{r^2 + v^2},
  \end{equation}
  visualized in Fig.~\ref{fig:nucleon_schematic}, which approximates the rms radius of the nucleon when the number of sampled constituents is large.
  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} rms radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence once should account for these difference when discussing the nucleon's rms radius.
  We account for this effect when discussing the nucleon width posterior distribution later in the text.

@@

\begin{table*}
  \caption{
    \label{tab:observables}
    Experimental data to be compared with model calculations.
  }
  \begin{ruledtabular}
  \begin{tabular}{lcccc}
    Collision system & Observable & Particle selection & Event bins & Ref. \\
    \paddedhline
    \multirow{2}{*}{Pb-Pb 5.02 TeV} & Yield $d\nch/d\eta$ & chg \& $|\eta| < 0.5$ & centrality: 0--5, 5--10, 10--20, \ldots, 60--70 & \note \\
    \noalign{\smallskip}
    & Flow cumulants $\vnk{n}{2}$ & chg \& $|\eta| < 1$ & centrality: 0--5, 5--10, 10--20, \ldots, 60--70 & \note \\
    \noalign{\smallskip}
    %Mean transverse momentum $\langle p_T \rangle$ & $\pi^\pm$, $K^\pm$, $p\bar p$ &
  \end{tabular}
  \end{ruledtabular}
\end{table*}

@@

\subsection{Calibration observables}

We focus on simple nuclear collision observables which are sensitive to the bulk properties of the produced medium and which can be calculated using a relatively small $\mathcal{O}(10^4)$ number of events.

>>>>

\usepackage{multirow}

@@

\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
\newcommand{\nch}{N_\text{ch}}
\newcommand{\ntrk}{N_\text{trk}^\text{offline}}
\newcommand{\sqrts}{\sqrt{s_{NN}}}
\newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
\newcommand{\X}{\chi_\mathrm{struct}}
\newcommand{\taufs}{\tau_\mathrm{fs}}
\newcommand{\dmin}{d_\mathrm{min}}
\newcommand{\T}{\tilde{T}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\vnk}[2]{v_#1\{#2\}}
\newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
\newcommand{\order}[1]{$\mathcal O(10^{#1})$}
\newcommand{\note}{\textcolor{theblue}{[?]}}

\renewcommand\labelitemi{$\vcenter{\hbox{\scriptsize$\bullet$}}$}

@@

Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass number and set of Woods-Saxon coefficients \cite{MOLLER1995185, DEVRIES1987495}, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to simulation predictions, outweighed by large uncertainties in modeling initial energy deposition and off-equilibrium dynamics \cite{Niemi:2014lha, Song:2011hk, Retinskaya:2013gca, Liu:2015nwa, Kurkela:2016vts}.
The geometry of light ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ signficantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
These nucleon substructure properties are difficult to measure and calculate from first principles and hence contribute significant uncertainty to model predictions of small systems.

@@

Within a picture of independent pairwise collisions between the constituents, a Glauber model model may be used to calculate the probability $P_\mathrm{coll}$ that the two protons collide inelastically at impact parameter $b$. In the limit when the number of constituents is large, it yields the particularly simple form
\begin{equation}
  \label{eq:pcoll}
  P_\mathrm{coll} = 1 - \exp[-\sigma_\mathrm{eff}\, T_{pp}(b)], \\
\end{equation}
where $\sigma_\mathrm{eff}$ is an effective cross section for pairwise inelastic collisions between the constituents, and $\sigma_{pp}^\mathrm{inel}$ is the total inelastic proton-proton cross section:
\begin{equation}
  \label{eq:sigma_nn}
  \sigma_{pp}^\mathrm{inel} = \int 2 \pi b\, db\, P_\mathrm{coll}(b).
\end{equation}
The proton densities $\rho_{A,B}$ in Eqn.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eqn.~\eqref{eq:pcoll}.
Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the proton size is comparable to the size of the produced QGP medium.

@@

In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom are uncertain as depicted in Fig.~\ref{fig:substructure}.
We model each nucleon's density $\rho_{A,B}$ as a sum of $n_c$ independent constituents
\begin{equation}
  \label{eq:rho}
  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\mathbf{x} - \mathbf{x_i}),
\end{equation}
where each constituent density $\rho_c$ is described by a Gaussian distribution of width $v$
\begin{equation}
  \label{eq:constituent_density}
  \rho_c(\mathbf{x}) = \frac{1}{(2\pi v^2)^{3/2}} \exp\left(-\frac{\x^2}{2 v^2}\right),
\end{equation}
and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution with standard deviation $r$.

@@

A simple remedy is to replace the arithmetic mean in the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
\end{equation}
based on a family of functions known as the generalized mean(s).
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto
  \begin{cases}
    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(T_A, T_B) & p \rightarrow -\infty.
  \end{cases}
\end{equation}
Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.

@@

We thus follow the procedure of the authors and reinterpret our initial transverse entropy density at midrapidity $s(\mathbf{x_\perp}, \tau_0)\vert_{\eta=0}$ as a transverse density of noninteracting massless partons $n(\mathbf{x}, \tau_0)\vert_{\eta=0}$.
The free-streamed energy-momentum tensor $T^{\mu\nu}(x, y, \eta, \tau)$ at transverse coordinate $(x, y)$, midrapidity $\eta=0$, and proper time $\tau > \tau_0$ is then given by
\begin{multline}
  \label{energy-momentum}
  T^{\mu\nu}(x, y, 0, \tau) = \frac{1}{\tau} \int_{0}^{2\pi} d\phi\, n(x - \Delta \tau \cos \phi, y - \Delta \tau \sin \phi) \\ \times
  \begin{bmatrix}
    1 & \cos\phi & \sin\phi  & 0\\
    \cos\phi & \cos^2\phi & \cos\phi\sin\phi & 0 \\
    \sin\phi & \sin\phi \cos\phi & \sin^2\phi & 0\\
    0 & 0 & 0 & 0
  \end{bmatrix},
\end{multline}
where $\Delta\tau$ is the free-streaming time $\Delta\tau = \tau - \tau_0$.
The solution \eqref{energy-momentum} is decomposed in hydrodynamic form
\begin{equation}
  \label{hydro-eqn}
  T^{\mu\nu} = e u^\mu u^\nu - (P + \Pi) \Delta^{\mu\nu} + \pi^{\mu\nu},
\end{equation}
where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} \equiv g^{\mu\nu} - u^\mu u^\nu}$ is the projector onto the space orthogonal to $u^\mu$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue,
\begin{equation}
  T^{\mu\nu} u_\nu = e u^\mu.
\end{equation}
The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eqn.~\eqref{hydro-eqn}
\begin{align}
  \Pi &= -\frac{1}{3} \mathrm{Tr}(\Delta_{\mu\nu} T^{\mu\nu}) - P,\\
  \pi^{\mu\nu} &= T^{\mu\nu} - e u^\mu u^\nu + (P + \Pi) \Delta^{\mu\nu}.
\end{align}

@@

The hydrodynamic equations of motion are necessarily closed using an equation of state (EoS) to relate the energy density $e$ and pressure $P$ of the produced medium.
We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval
$165 \le T \le 200$~MeV \cite{Bernhard:2018hnz}.
For the lattice EoS, we use a calculation by the HotQCD collaboration for (2+1)-flavor QCD which was extrapolated to the continuum limit \cite{Bazavov:2014pvz}. Recent developments in lattice QCD now enable calculations in (2+1+1)-flavors \cite{Borsanyi:2016ksw}, i.e.\ with thermalized charm quarks, and the additional charm flavor has been shown to visibly affect predictions of $p_T$-differential flow observables \cite{Noronha-Hostler:2018zxc}.
Investigating this sensitivity would thus be a natural target for future improvements to the present work.

@@

For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Tbl.~\ref{tab:design}.
We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
Several of the model parameters require special care and are reparametrized accordingly:
\begin{itemize}[leftmargin=2\parindent]
  \item The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
    The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
    We therefore reparametrize the constituent fluctuations using the standard deviation of the resulting nucleon fluctuations
    \begin{equation}
      \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
    \end{equation}
    where $k$ is the shape parameter of the gamma distribution used to fluctuate each constituent.
  \item In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
    \begin{enumerate*}[label=(\roman*)]
      \item a parameter $v$ for the Gaussian width of each constituent,
      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
    \end{enumerate*}
    In the limit $n_c \to \infty$, the composite nucleon's root mean square (rms) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
  \begin{equation}
    \lim_{n_c \to\infty} \mathrm{RMS}\{\rho(\mathbf{x})\} = \sqrt{r^2 + v^2}.
  \end{equation}
  We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
  \begin{equation}
    \label{eq:nucleon_width}
    w = \sqrt{r^2 + v^2},
  \end{equation}
  depicted in Fig.~\ref{fig:nucleon_schematic}, which approximates the rms radius of the nucleon when the number of sampled constituents is large.
  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} rms radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior rms radius.

@@

\begin{table*}
  \caption{
    \label{tab:observables}
    Experimental data used to calibrate the model.
  }
  \begin{ruledtabular}
  \begin{tabular}{cc}
    Pb-Pb $\sqrts=5.02$~TeV & $p$-Pb $\sqrts=5.02$~TeV \\
    \paddedhline
    Charged-particle multiplicity $d\nch/d\eta$, $|\eta| < 0.5$ \cite{Adam:2015ptt} & Charged-particle multiplicity $d\nch/d\eta$, $|\eta| < 1.4$ \cite{Adam:2014qja} \\
    \noalign{\smallskip}
  Two-particle flow cumulants  $\vnk{n}{2}$ for $n=2,3,4$, $|\eta| < 0.8$,  & Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3$, $|\eta| < 2.4$, \\
    charged-particles, $|\Delta\eta| > 1$,\, $0.2 < p_T < 5.0$~GeV \cite{Adam:2016izf} & charged-particles, $|\Delta\eta| > 2$,\, $0.3 < p_T < 3.0$~GeV \cite{Chatrchyan:2013nka}\\
    \noalign{\smallskip}
    & Charged-particle mean $p_T$, $0.15 < p_T < 10$~GeV, $|\eta| < 0.3$ \cite{Abelev:2013bla}\\
  \end{tabular}
  \end{ruledtabular}
\end{table*}

@@

\subsection{Simulated and measured observables}

The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\x$ by comparing the model predictions $\y_m(\x)$ to experimental data $\y_e$.
We focus on simple experimental observables in the present study which are sensitive to the bulk properties of the produced medium.
We calculate for each set of model parameters the following observables at midrapidity:
\begin{itemize}[leftmargin=2\parindent, itemsep=0pt]
  \item Charged-particle multiplicity $d\nch/d\eta$ \cite{Adam:2015ptt, Adam:2014qja}.
  \item Identified particle yields $dN/dy$ of pions, kaons, and protons.
  \item Transverse energy production $dE_T/d\eta$.
  \item Charged particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV) \cite{Abelev:2013bla}.
  \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons and protons.
  \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged particles, $0.15 < p_T < 2.0$~GeV).
  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV \cite{Adam:2016izf}, and $0.3 < p_T < 3.0$~GeV \cite{Chatrchyan:2013nka}).
\item Four-particle flow cumulant $\vnk{2}{4}$\\ (charged particles, $0.2 < p_T < 5.0$) \cite{Adam:2016izf}.
  \item Symmetric cumulants $\mathrm{SC}(4, 2)$ and $\mathrm{SC}(3,2)$.
\end{itemize}
Each observable is calculated from the list of final state particles produced by \texttt{UrQMD} using the same methods applied by experiment.
We generally match the kinematic cuts of all measurements with two exceptions: we use use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a finite rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.

At the time of this writing, many of the aforementioned experimental observables are not yet published for $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
We therefore restrict our calibration to the subset of measured and published observables listed in Tbl.~\ref{tab:observables}.
The Pb-Pb four-particle cumulants $\vnk{n}{4}$ at $\sqrts=5.02$~TeV which are noticeably absent \emph{are} currently available, but we exclude from the present calibration for statistical reasons.

Most of the calibration observables listed in Tbl.~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined as some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
When calculating these observables, we generate $\mathcal{O}(10^4)$ minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.

However, for some observables such as $p$-Pb mean $p_T$ and flow cumulants, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline \note.
These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
We exploit, for this purpose, the approximate correspondence
\begin{equation}
  \ntrk \sim \nch \propto \int d^2x\, n(\x)
\end{equation}
between each event's initial transverse parton density $n(\x)$ and its final-state event activity, characterized by $\nch$ or $\ntrk$.

and the event's initial parton density $n(\x)$ in Eqn.~\eqref{energy-momentum}, and run a high-multiplicity trigger on the relative event activity
\begin{equation}
  \mathrm{bin~low} \le \nch / \langle \nch \rangle \le \mathrm{bin~high}.
\end{equation}

\fig*{correlation_matrices}{
  Correlation matrices for Pb-Pb collisions at $\sqrts=5.02$~TeV for the model/emulator (left) and experimental data (right) used to construct the covariance matrix $\Sigma = \Sigma_\mathrm{model} + \Sigma_\mathrm{expt}$ in Eqn.~\eqref{eq:likelihood}.
  Experimental statistical and systematic errors are from ALICE and CMS \note.
  The experimental correlation structure is modeled using Eqn.~?.
}
